Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-05 latex/robdml.tex
tex

Although attacks against DML has been widely studied~\cite{advrank,advorder},
the defense methods for improving adversarial robustness is much less explored.
%
As adversarial training~\cite{madry} remains to be one of the most effective
defense for classification, existing defense methods for DML are also
adversarial training-based.
%
Embedding-Shifted Triplet (EST)~\cite{advrank} adopts adversarial counterpars
of $\va,\vp,\vn$ with maximum embedding move distance off their original
locations, \ie,
$L_\text{EST}=L_\text{trip}(\tilde{\va},\tilde{\vp},\tilde{\vn};\gamma)$ where
$\tilde{\va}=\phi(X+r^*)$, $r^*=\arg\max_{r}d_\phi(X+r, X)$.
%
Anti-Collapse Triplet (ACT)~\cite{robrank} collapses the embedding vectors
of positive and negative sample, and enforces the model to separate them
apart, \ie, $L_\text{ACT}=L_\text{trip}(\va, \overrightarrow{\vp},
\overleftarrow{\vn};\gamma)$, where $[\overrightarrow{\vp},\overleftarrow{\vn}]
=[\phi(X_p+r_p^*), \phi(X_n+r_n^*)]$, and $[r_p^*,r_n^*]=\arg\min_{r_p,r_n}
d_\phi(X_p+r_p, X_n+r_n)$.
%
\oo{Their issues.}

@@

>>>>

Although attacks against DML has been widely studied~\cite{advrank,advorder},
the defense methods for improving adversarial robustness is much less explored.
%
As adversarial training~\cite{madry} remains to be one of the most effective
defense for classification, existing defense methods for DML are also
adversarial training-based.
%
Embedding-Shifted Triplet (EST)~\cite{advrank} adopts adversarial counterpars
of $\va,\vp,\vn$ with maximum embedding move distance off their original
locations, \ie,
$L_\text{EST}=L_\text{trip}(\tilde{\va},\tilde{\vp},\tilde{\vn};\gamma)$ where
$\tilde{\va}=\phi(X+r^*)$, $r^*=\arg\max_{r}d_\phi(X+r, X)$.
%
Anti-Collapse Triplet (ACT)~\cite{robrank} collapses the embedding vectors of
positive and negative sample, and enforces the model to separate them apart,
\ie, $L_\text{ACT}=L_\text{trip}(\va, \overrightarrow{\vp},
\overleftarrow{\vn};\gamma)$, where $[\overrightarrow{\vp},\overleftarrow{\vn}]
=[\phi(X_p+r_p^*), \phi(X_n+r_n^*)]$, and $[r_p^*,r_n^*]=\arg\min_{r_p,r_n}
d_\phi(X_p+r_p, X_n+r_n)$.
%
However, compared to the standard min-max adversarial training
paradigm~\cite{madry}, these methods merely indirectly increase the loss value,
and thus, suffer from inefficient learning because the adversary is not strong
enough.
%
\oo{Furthermore, min-max form -- free adversarial training}.

@@

Given an anchor image $X_a$, we sample a positive image $X_p^S$ (in the same
class as the anchor) and a negative image $X_n^S$ (in different class as the
anchor) with a certain sampling strategy (\eg, semi-hard~\cite{facenet}).
%
For convenience we call this triplet as ``source triplet''.
%
Then its \emph{hardness} is defined as
$H(X_a,X_p^S,X_n^S)=d_\phi(X_a,X_p^S)-d_\phi(X_a,X_n^S)$, which is an internal
part of the triplet loss.

In the traditional min-max adversarial training setting~\cite{madry}, the inner
``max'' problem should maximize $H(X_a,X_p^S,X_n^S)$ and hence maximize the triplet
loss.
%
However, existing defense methods~\cite{advorder,robrank} refrain from adopting
such paradigm because the model will quickly collapse with excessively hard
triplets~\cite{facenet}, and adopt relatively weak adversaries instead
(inefficient in gaining robustness).
